Gut microbes and their genes or something else

### 2023-02-17 manuscript/tex/sections/discussion.tex
citations in discussion

The use of shotgun metagenomic sequencing enabled us to get
species-level resolution of microbial taxa. A previous study of
cognition in 3 year old subjects used 16S rRNA gene amplicon sequencing,
and showed that genera from the Lachnospiraceae family as well as
unclassified Clostridiales (now Eubacteriales) were associated with
higher scores on the Ages and Stages Questionnaire
\href{https://www.zotero.org/google-docs/?SB5Lwa}{(Sordillo et al.,
2019)}. However, each of these clades encompass dozens of genera with
diverse functions, each of which may have different effects. Indeed,
several of the taxa that were positively associated with cognitive
function in this study, including \emph{B. wexlerae}, the only species
identified by linear models in children under 6 months, \emph{D.
longicatena}, \emph{R. faecis}, and \emph{A. finegoldii} are
Clostridiales, as is \emph{R. gnavus}, which we found was negatively
associated with cognitive function (Figure 2A-B). This kind of
species-level resolution is typically not possible with amplicon
sequencing.

We identified several species in the family Eggerthelaceae that were
associated with cognitive function, including \emph{Gordonibacter
pamelaeae, Aldercreutzia equolofaciens, Asaccharobacter celatus}
(formerly regarded a subspecies of \emph{A. equolofaciens}
\href{https://www.zotero.org/google-docs/?OEhguI}{(Takahashi et al.,
2021)}), and \emph{Eggerthella lenta}. Many members of this family are
known in part due to unique metabolic activities. For example, \emph{A.
equolofaciens} produces the nonsteroidal estrogen equol from isoflaven
found in soybeans
\href{https://www.zotero.org/google-docs/?I2qNyB}{(Wang et al., 2005)},
and \emph{G. pamelaeae} can metabolize the polyphenol ellagic acid
(found in pomegranates and some berries) into urolithin, which has been
shown in some studies to have a neuroprotective effect
\href{https://www.zotero.org/google-docs/?oTN8Lw}{(Gong et al., 2022;
Selma et al., 2014)}. \emph{E. lenta} has been extensively studied for
its ability to metabolize drug compounds such as the plant-derived heart
medication digoxin
\href{https://www.zotero.org/google-docs/?tH2Qeq}{(Haiser et al.,
2013)}. The metabolic versatility of this clade, and the large number of
species that are associated with cognition make these microbes prime
targets for further mechanistic studies.

@@

The use of multiple age-appropriate cognitive assessments that could be
normalized to a common scale enabled us to analyze microbial
associations across multiple developmental periods, but carries several
drawbacks. In particular, the test-retest reliability, as well as
systematic differences between test administrators may introduce
substantial noise into these observations, particularly in the youngest
children. In addition, our study period overlapped with the beginning of
the COVID-19 pandemic, and we and others have observed some reduction in
measured scores for children that were assessed after the implementation
of lockdowns. In our subject set for this study, these effects are more
pronounced in some age groups due to our sampling schedule
\href{https://www.zotero.org/google-docs/?L8GnRj}{(Blackwell et al.,
2022; Deoni et al., 2021)} (Supplementary Figure 3).

>>>>

The use of shotgun metagenomic sequencing enabled us to get
species-level resolution of microbial taxa. A previous study of
cognition in 3 year old subjects used 16S rRNA gene amplicon sequencing,
and showed that genera from the Lachnospiraceae family as well as
unclassified Clostridiales (now Eubacteriales) were associated with
higher scores on the Ages and Stages Questionnaire
\cite{sordilloAssociationInfantGut2019}.
However, each of these clades encompass dozens of genera with
diverse functions, each of which may have different effects. Indeed,
several of the taxa that were positively associated with cognitive
function in this study, including \emph{B. wexlerae}, the only species
identified by linear models in children under 6 months, \emph{D.
longicatena}, \emph{R. faecis}, and \emph{A. finegoldii} are
Clostridiales, as is \emph{R. gnavus}, which we found was negatively
associated with cognitive function (Figure 2A-B). This kind of
species-level resolution is typically not possible with amplicon
sequencing.

We identified several species in the family Eggerthelaceae that were
associated with cognitive function, including \emph{Gordonibacter
pamelaeae, Aldercreutzia equolofaciens, Asaccharobacter celatus}
(formerly regarded a subspecies of \emph{A. equolofaciens}
\cite{takahashiCompleteGenomeSequence2021}),
and \emph{Eggerthella lenta}. Many members of this family are
known in part due to unique metabolic activities. For example, \emph{A.
equolofaciens} produces the nonsteroidal estrogen equol from isoflaven
found in soybeans
\cite{wangEnantioselectiveSynthesisSEquol2005},
and \emph{G. pamelaeae} can metabolize the polyphenol ellagic acid
(found in pomegranates and some berries) into urolithin, which has been
shown in some studies to have a neuroprotective effect
\cite{gongUrolithinAlleviatesBloodbrain2022,selmaDescriptionUrolithinProduction2014}.
\emph{E. lenta} has been extensively studied for
its ability to metabolize drug compounds such as the plant-derived heart
medication digoxin
\cite{haiserPredictingManipulatingCardiac2013}.
The metabolic versatility of this clade, and the large number of
species that are associated with cognition make these microbes prime
targets for further mechanistic studies.

@@

The use of multiple age-appropriate cognitive assessments that could be
normalized to a common scale enabled us to analyze microbial
associations across multiple developmental periods, but carries several
drawbacks. In particular, the test-retest reliability, as well as
systematic differences between test administrators may introduce
substantial noise into these observations, particularly in the youngest
children. In addition, our study period overlapped with the beginning of
the COVID-19 pandemic, and we and others have observed some reduction in
measured scores for children that were assessed after the implementation
of lockdowns. In our subject set for this study, these effects are more
pronounced in some age groups due to our sampling schedule
\cite{blackwellYouthWellbeingCOVID192022,deoniImpactCOVID19Pandemic2021}
(Supplementary Figure 3).
